---
title: "Working with MySQL in R"
author: "Illarion  Jabine"
date: "19/01/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Before you begin you might need to execute this SQL in MySQL:

ALTER USER 'username'@'localhost'
IDENTIFIED WITH mysql_native_password BY 'password'


```{r MySQL connection}
library(RMySQL)
db_user <- 'root'
db_password <- 'test'
db_name <- 'test'
db_table <- 'test_table1'
db_host <- '127.0.0.1' # for local access
db_port <- 3306
mydb <-  dbConnect(MySQL(), user = db_user, password = db_password,
                   dbname = db_name, host = db_host, port = db_port)
```

```{r SQL: create table, insert values and select}
# SQL create table
create_sql <- paste0("CREATE TABLE ",paste0(db_table)," (`test_column` INT NULL)")
dbSendQuery(mydb, create_sql)

# SQL insert
for (i in 1:100) {
insert_sql <- paste0("insert into ",paste0(db_table)," values(",paste0(i),")")
dbGetQuery(mydb, insert_sql)
}

# SQL select
select_sql <- paste0("select * from ", db_table)
res <- dbSendQuery(mydb, select_sql)
df <-  fetch(res, n = -1)
on.exit(dbDisconnect(mydb))

```

With dbWriteTable() we can write an R dataframe directly into MySQL.
This function will automatically create an SQL tabl definition and insert the value.
Make sure that you enable "local-infile" parameter in "C:\ProgramData\MySQL\MySQL Server 8.0\my.ini" file 
[mysqld]
local-infile=1

```{r writing a whole dataframe}
# Write a whole dataframe into db
dbWriteTable(mydb, "mtcars", mtcars)
dbReadTable(mydb, "mtcars")
```




